Information Distribution Depends on Language-Specific Features

### 2019-02-02 cogsci/position-info.Rmd
final cogsci push

Over 7,000 languages are spoken around the modern world [@simons2018]. These language vary along many dimensions, but all share a core goal: communicating information. If speakers and writers of these languages act near-optimally to achieve their communicative goals, regularities of use across these diverse languages can be explained by a rational theory of communication [@anderson1991]. 
Information theory, a mathematical framework developed by @shannon1948 to describe the transmission and decoding of signals, has been a unifying approach for the recent development of theories of communication in human and machine language processing [@jelinek1976;@levy2007].

@@

The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade. The original evidence in favor of Uniform Information Density in @levy2007 finds that the insertion of complementizers (e.g. "that") in relative clauses in English corresponds to where neighboring words have high information content. Similarly, @frank2008 argues that contractions in English such as "you're" do not occur when neighboring words are highly informative. Applications of Uniform Information Density include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.

However, other recent work has contradicted the Uniform Information Density hypothesis. Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in sentences. @zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high. @jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi speaker's word order.

@@

We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in English--in written language, and in conversational speech between adults and between parents and their children. Thus, this entropy curve is a robust feature of English linguistic productions. We then examine entropy curves cross-linguistically, and show that characteristic curves vary across the world's languages. Finally, we show that this variation is predictable in part from the structure of individual languages (i.e. word order). Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: speakers may structure their utterances to optimize information density, but they must do so under the predictable constraints of their language.

@@

With this measure, we can compute the unigram entropy at each word position of sentences of each length within a corpus. The result of this method can be plotted for each sentence length as an *entropy curve*, which can be visually compared across sentence lengths to observe how the unigram entropy changes across absolute positions in each of the sentences. @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus. Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech [@levy2007], and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing affine function at the word level.

@@

We began by replicating Yu et al.'s [-@yu2016] analysis, computing entropy curves on the British National Corpus--a collection of predominantly written English [@clear1993]. We then applied the same method to the Switchboard corpus--a collection of spoken English [@godfrey1992]. This allowed us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of written language.

# Data and Analysis

The British National Corpus consists of predominantly (90%) written documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc). It also contains a small collection of corpora from spoken language. All together, the British National Corpus contains $\sim$ 100 million tokens. The Switchboard corpus is a collection of $\sim$ 2,400 telephone conversations between unacquainted adults prompted with subjects of conversation. Switchboard is the corpus used in Levy and Jaeger's [-@levy2007] original demonstration of the Uniform Information Density Hypothesis.

@@

# Results

The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lengths we analyzed (Figure \ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s [-@yu2016] findings, we find that
the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again for the sentence-final word.

```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}

@@

bnc_plot <- ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length, scales = "free_x") + 
  xlab("British National Corpus utterance position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25) +
  geom_line(size = .25) +
  theme_classic(base_size = 10)

sw_plot <- ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length, scales = "free_x") + 
  xlab("Switchboard utterance position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25) +
  geom_line(size = .25) +
  theme_classic(base_size = 10)

@@

ggplot(childes_entropies, aes(x = word_order, y = empirical_entropy,
                      color = person,
                      label = person)) +
  facet_grid(language ~ length, scales = "free") + 
  xlab("Utterance position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_line(size = .25) +
  theme_classic(base_size = 10) + 
  scale_color_ptol() +
  theme(legend.position = "nones") + 
  geom_text(data = label_data)
```

@@

Wikipedia has two primary advantages: (1) Individual language corpora are generally large, and (2) hundreds of languages are represented. To understand how variability in entropy curves is related to the structure of language, we used features from the World Atlas of Language Structures [WALS, @dryer2013]. WALS is a curated collection of typological features which analyze and classify the diversity across the world's languages. However, as the collection includes contributions from many authors studying a heterogeneous set of topics and languages, most features are missing for most languages. For our sample, we sought to balance diversity of languages with coverage in WALS. We selected $45$ languages for which eight specific WALS features were coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all reflected the order of words in a sentence in each language (e.g. whether objects come before verbs, whether adjectives come before nouns). A consistent set of results across these features would be consistent with an effect of order on entropy curves, but unfortunately no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In order to understand how these curves varied across languages, we developed a method for compressing the curves into a few representative values. We noticed that curves tended to vary in predictable utterance positions: The majority were flat in the middle of utterances, but had either rising or falling slopes at the beginnings and ends. For each language, for each utterance length $k$ we computed 5 slopes: The change between the 1st and 2nd position, the change between the 2nd and 3rd position, the slope between the 3rd position and the 3rd-to-last position ($k-1$), the slope between the 3rd-to-last and the 2nd-to-last position ($k-1$), and the slope between the 2nd-to-last and final positions. We then averaged these slopes across utterance lengths within a language to estimate a language-typical signature. We performed an additional analysis dividing each utterance into fifths evenly and computing the slopes between these fifths. Results from this second analysis were qualitatively similar.

@@

We treated these 5 slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997]. Intuitively, vectors with a smaller angle between them are more similar. Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly, e.g. Slavic languages are similar to each-other.

@@

We then asked whether these pairwise similarities were related to overlap in linguistic features. For each of the 8 WALS features we measured, we fit a logistic regression predicting whether two languages would have the same value for that feature from the similarity of their pairwise entropy curves. Overlap on all 8 features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}. Order of object and verb ($83A$) was the strongest predictor, explaining `r max_r2`% of the variance in pairwise entropy curve similarity.

@@

We expect the characteristic entropy curve for a language to have important downstream effects for how people process that language. While conversational turn-taking is universal, there exists cross-linguistic variation in the average length of time between conversational partners speaking [@stivers2009universals]. We predict that one factor in this variability relates to the shape of a language's entropy curve. A jump at the end of a language's entropy curve, such as in English, indicates that there is on average high levels of information at the end of utterances in that language. Therefore we expect turn-taking times to be longer in such languages than in languages where there is a drop in the final segment of the entropy curve, such as Mandarin. We also expect that readers of a language will take longer on average to read parts of sentences where the entropy curve is relatively high, due to the presence of higher information content. This may explain the so-called wrap-up effect in sentence processing, in which the end of the sentence in English and other Germanic languages is processed more slowly during reading [@stowe2018;@kuperman2010].

>>>>

Over 7,000 languages are spoken around the modern world [@simons2018]. These languages vary along many dimensions, but all share a core goal: communicating information. If speakers and writers of these languages act near-optimally to achieve their communicative goals, regularities of use across these diverse languages can be explained by a rational theory of communication [@anderson1991]. 
Information theory, a mathematical framework developed by @shannon1948 to describe the transmission and decoding of signals, has been a unifying approach for the recent development of theories of communication in human and machine language processing [@jelinek1976;@levy2007].

@@

The evidence in favor of Uniform Information Density has largely focused on specific  points within sentences, while the hypothesis itself has been applied broadly over the past decade. The original evidence in favor of Uniform Information Density in @levy2007 finds that the insertion of complementizers (e.g. "that") in relative clauses in English corresponds to where neighboring words have high information content. Similarly, @frank2008 argues that contractions in English such as "you're" do not occur when neighboring words are highly informative. Applications of Uniform Information Density include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.

However, other recent work has contradicted the Uniform Information Density hypothesis. Similar to the original work in @levy2007, @zhan2018 focuses on information distribution at particular points in sentences. @zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high. Examining the information density at the sentence level, @jain2018 study word order across written sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi writer's word order.

@@

We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found throughout English--in written language, and in conversational speech between adults and between parents and their children. Thus, this entropy curve is a robust feature of English linguistic productions. We then examine entropy curves cross-linguistically, and show that the characteristic entropy curves vary across the world's languages. Finally, we show that this variation is predictable in part from the structure of individual languages (i.e. word order). Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: speakers may structure their utterances to optimize information density, but they must do so under the predictable constraints of their language.

@@

With this measure, we can compute the unigram positional entropy at each word position of sentences of each length within a corpus. The result of this method across word positions can be plotted for each sentence length as an *entropy curve*, which can be visually compared across sentence lengths to observe how the unigram entropy changes across absolute positions in each of the sentences. @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus. Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech [@levy2007], and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing affine function at the word level.

@@

We began by replicating Yu et al.'s [-@yu2016] analysis, computing entropy curves on the British National Corpus--a collection of predominantly written English [@clear1993]. We then applied the same method to the Switchboard corpus--a collection of spoken English [@godfrey1992]. This allowed us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of their written language corpus.

The British National Corpus consists of predominantly (90%) written documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc). It also contains a small collection of corpora from spoken language. All together, the British National Corpus contains $\sim$ 100 million tokens. The Switchboard corpus is a collection of $\sim$ 2,400 telephone conversations between unacquainted adults prompted with subjects of conversation. Switchboard is the corpus used in Levy and Jaeger's [-@levy2007] original demonstration of the Uniform Information Density hypothesis.

@@

The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lengths we analyzed (Figure \ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s [-@yu2016] findings, we found that
the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again for the sentence-final word.

```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points represent average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}

@@

bnc_plot <- ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length, scales = "free_x") + 
  xlab("British National Corpus word position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25) +
  geom_line(size = .25) +
  theme_classic(base_size = 10)

sw_plot <- ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length, scales = "free_x") + 
  xlab("Switchboard word position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25) +
  geom_line(size = .25) +
  theme_classic(base_size = 10)

@@

ggplot(childes_entropies, aes(x = word_order, y = empirical_entropy,
                      color = person,
                      label = person)) +
  facet_grid(language ~ length, scales = "free") + 
  xlab("Word position") +
  ylab("Positional entropy") +
  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_line(size = .25) +
  theme_classic(base_size = 10) + 
  scale_color_ptol() +
  theme(legend.position = "nones") + 
  geom_text(data = label_data)
```

@@

Wikipedia has two primary advantages: (1) Individual language corpora are generally large, and (2) hundreds of languages are represented. To understand how variability in entropy curves is related to the structure of language, we used features from the World Atlas of Language Structures [WALS, @dryer2013]. WALS is a curated collection of typological features which help to classify the diversity across the world's languages. However, as the collection includes contributions from many authors studying a heterogeneous set of topics and languages, most features are missing for most languages. For our sample, we sought to balance diversity of languages with coverage in WALS. We selected eight specific WALS features for which the same $45$ languages were coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all reflected the order of words in a sentence in each language (e.g. whether objects come before verbs, whether adjectives come before nouns). A consistent set of results across these features would be consistent with an effect of order on entropy curves, but unfortunately no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In order to understand how these curves varied across languages, we developed a method for compressing the curves into a few representative values. We noticed that curves tended to vary at predictable positions within sentences: The majority were flat in the middle, but had either rising or falling slopes at the beginnings and ends. For each language, for each sentence length $k$ we computed 5 slopes: The change between the 1st and 2nd position, the change between the 2nd and 3rd position, the slope between the 3rd position and the 3rd-to-last position ($k-1$), the slope between the 3rd-to-last and the 2nd-to-last position ($k-1$), and the slope between the 2nd-to-last and final positions. We then averaged these slopes across sentence lengths within a language to estimate a language-typical signature. We performed an additional analysis dividing each sentence into fifths evenly and computing the slopes between these fifths. Results from this second analysis were qualitatively similar.

@@

We treated these five slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997]. Intuitively, vectors with a smaller angle between them are more similar. Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly (e.g. Slavic languages are similar to each other).

@@

We then asked whether these pairwise similarities were related to overlap in linguistic features. For each of the eight WALS features we measured, we fit a logistic regression predicting whether two languages would have the same value for that feature from the pairwise similarity of their entropy curves. Overlap on all eight features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}. Order of object and verb ($83A$) was the strongest predictor, explaining `r max_r2`% of the variance in pairwise entropy curve similarity.

@@

We expect the characteristic entropy curve for a language to have important downstream effects for how people process that language. While conversational turn-taking is universal, there exists cross-linguistic variation in the average length of time between conversational partners speaking [@stivers2009universals]. We predict that one factor in this variability relates to the shape of a language's entropy curve. A jump at the end of a language's entropy curve, such as in English, indicates that there is on average high levels of information at the end of utterances in that language. Therefore we expect turn-taking times to be longer in such languages than in languages where there is a drop in the final segment of the entropy curve, such as Mandarin. We also expect that readers of a language will take longer on average to read parts of sentences where the entropy curve is relatively high, due to the presence of higher information content. This may explain the so-called wrap-up effect in sentence processing, in which the ends of sentences are processed more slowly during reading [@stowe2018;@kuperman2010]. The entropy curve analysis can be carried out on any language across a wide variety of text corpora, making it a powerful tool for analyzing the information structure of both spoken and written utterances in a language.

### 2019-02-02 cogsci/position-info.tex
final cogsci push

Over 7,000 languages are spoken around the modern world (Simons \&
Fenning, 2018). These language vary along many dimensions, but all share
a core goal: communicating information. If speakers and writers of these
languages act near-optimally to achieve their communicative goals,
regularities of use across these diverse languages can be explained by a
rational theory of communication (Anderson, 1991). Information theory, a
mathematical framework developed by Shannon (1948) to describe the
transmission and decoding of signals, has been a unifying approach for
the recent development of theories of communication in human and machine
language processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).

@@

The evidence in favor of Uniform Information Density largely been
situation-specific and English-language driven, while the hypothesis
itself has been applied broadly over the past decade. The original
evidence in favor of Uniform Information Density in R. P. Levy \& Jaeger
(2007) finds that the insertion of complementizers (e.g. ``that'') in
relative clauses in English corresponds to where neighboring words have
high information content. Similarly, A. F. Frank \& Jaeger (2008) argues
that contractions in English such as ``you're'' do not occur when
neighboring words are highly informative. Applications of Uniform
Information Density include determining whether linguistic alignment
takes place (Jaeger \& Snider, 2013), Zipfian word length distributions
(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
Reitter, 2018) and the significance of ambiguity in language
(Piantadosi, Tily, \& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original R. P. Levy \& Jaeger (2007),
Zhan \& Levy (2018) focuses on information distribution at particular
points in sentences. Zhan \& Levy (2018) finds that more
information-rich classifiers in Mandarin Chinese are produced when
production of the neighboring noun is difficult, not when the
information content is high. Jain, Singh, Ranjan, Rajkumar, \& Agarwal
(2018) examine word order across spoken sentences in Hindi, a freer word
order language than English, and find that information density has no
significant effect on determining a Hindi speaker's word order.

@@

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found in English--in written language, and in
conversational speech between adults and between parents and their
children. Thus, this entropy curve is a robust feature of English
linguistic productions. We then examine entropy curves
cross-linguistically, and show that characteristic curves vary across
the world's languages. Finally, we show that this variation is
predictable in part from the structure of individual languages
(i.e.~word order). Taken together, our results suggest a refinement of
the Uniform Information Density hypothesis: speakers may structure their
utterances to optimize information density, but they must do so under
the predictable constraints of their language.

@@

With this measure, we can compute the unigram entropy at each word
position of sentences of each length within a corpus. The result of this
method can be plotted for each sentence length as an \emph{entropy
curve}, which can be visually compared across sentence lengths to
observe how the unigram entropy changes across absolute positions in
each of the sentences. Genzel \& Charniak (2002) similarly examine a
unigram entropy measure on sentences, and found that entropy at the
sentence level increases linearly with sentence index within a corpus.
Uniform Information Density applies this uniformity of entropy rate in
sentences to all levels of speech (R. P. Levy \& Jaeger, 2007), and so
our method obtained from Yu et al. (2016), which examines text at the
word level, should find a monotonically increasing affine function at
the word level.

@@

We began by replicating Yu et al.'s (2016) analysis, computing entropy
curves on the British National Corpus--a collection of predominantly
written English (Clear, 1993). We then applied the same method to the
Switchboard corpus--a collection of spoken English (Godfrey, Holliman,
\& McDaniel, 1992). This allowed us to ask whether the characteristic
function identified by Yu et al. (2016) is a general feature of English,
or instead a function of written language.

\section{Data and Analysis}\label{data-and-analysis}

The British National Corpus consists of predominantly (90\%) written
documents collected in the early 1980s and 1990s across a variety of
genres (scientific articles, newspapers, fiction, etc). It also contains
a small collection of corpora from spoken language. All together, the
British National Corpus contains \(\sim\) 100 million tokens. The
Switchboard corpus is a collection of \(\sim\) 2,400 telephone
conversations between unacquainted adults prompted with subjects of
conversation. Switchboard is the corpus used in Levy and Jaeger's (2007)
original demonstration of the Uniform Information Density Hypothesis.

@@

\section{Results}\label{results}

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lengths we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we find that the positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again for
the sentence-final word.

@@

\caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
\end{figure*}
\end{CodeChunk}

@@

Wikipedia has two primary advantages: (1) Individual language corpora
are generally large, and (2) hundreds of languages are represented. To
understand how variability in entropy curves is related to the structure
of language, we used features from the World Atlas of Language
Structures (WALS, Dryer \& Haspelmath, 2013). WALS is a curated
collection of typological features which analyze and classify the
diversity across the world's languages. However, as the collection
includes contributions from many authors studying a heterogeneous set of
topics and languages, most features are missing for most languages. For
our sample, we sought to balance diversity of languages with coverage in
WALS. We selected \(45\) languages for which eight specific WALS
features were coded (Figure \ref{fig:clust_tree}). Unfortunately, these
features all reflected the order of words in a sentence in each language
(e.g.~whether objects come before verbs, whether adjectives come before
nouns). A consistent set of results across these features would be
consistent with an effect of order on entropy curves, but unfortunately
no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In
order to understand how these curves varied across languages, we
developed a method for compressing the curves into a few representative
values. We noticed that curves tended to vary in predictable utterance
positions: The majority were flat in the middle of utterances, but had
either rising or falling slopes at the beginnings and ends. For each
language, for each utterance length \(k\) we computed 5 slopes: The
change between the 1st and 2nd position, the change between the 2nd and
3rd position, the slope between the 3rd position and the 3rd-to-last
position (\(k-1\)), the slope between the 3rd-to-last and the
2nd-to-last position (\(k-1\)), and the slope between the 2nd-to-last
and final positions. We then averaged these slopes across utterance
lengths within a language to estimate a language-typical signature. We
performed an additional analysis dividing each utterance into fifths
evenly and computing the slopes between these fifths. Results from this
second analysis were qualitatively similar.

@@

We treated these 5 slopes as a vector, allowing us to compute the
pairwise similarity between languages with cosine--a standard measure of
vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
a smaller angle between them are more similar. Figure
\ref{fig:clust_tree} shows a hierarchical clustering of these 5
dimensional vectors, demonstrating that entropy curve similarity appears
to capture typological relatedly, e.g.~Slavic languages are similar to
each-other.

We then asked whether these pairwise similarities were related to
overlap in linguistic features. For each of the 8 WALS features we
measured, we fit a logistic regression predicting whether two languages
would have the same value for that feature from the similarity of their
pairwise entropy curves. Overlap on all 8 features was reliably
predicted by entropy curve similarity (largest \(p = .001\)), with the
amount of variance in feature overlap predicted varying across
predictors \ref{tab:tab_features}. Order of object and verb (\(83A\))
was the strongest predictor, explaining 16.22\% of the variance in
pairwise entropy curve similarity.

@@

We expect the characteristic entropy curve for a language to have
important downstream effects for how people process that language. While
conversational turn-taking is universal, there exists cross-linguistic
variation in the average length of time between conversational partners
speaking (Stivers et al., 2009). We predict that one factor in this
variability relates to the shape of a language's entropy curve. A jump
at the end of a language's entropy curve, such as in English, indicates
that there is on average high levels of information at the end of
utterances in that language. Therefore we expect turn-taking times to be
longer in such languages than in languages where there is a drop in the
final segment of the entropy curve, such as Mandarin. We also expect
that readers of a language will take longer on average to read parts of
sentences where the entropy curve is relatively high, due to the
presence of higher information content. This may explain the so-called
wrap-up effect in sentence processing, in which the end of the sentence
in English and other Germanic languages is processed more slowly during
reading (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010; Stowe, Kaan,
Sabourin, \& Taylor, 2018).

>>>>

Over 7,000 languages are spoken around the modern world (Simons \&
Fenning, 2018). These languages vary along many dimensions, but all
share a core goal: communicating information. If speakers and writers of
these languages act near-optimally to achieve their communicative goals,
regularities of use across these diverse languages can be explained by a
rational theory of communication (Anderson, 1991). Information theory, a
mathematical framework developed by Shannon (1948) to describe the
transmission and decoding of signals, has been a unifying approach for
the recent development of theories of communication in human and machine
language processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).

@@

The evidence in favor of Uniform Information Density has largely focused
on specific points within sentences, while the hypothesis itself has
been applied broadly over the past decade. The original evidence in
favor of Uniform Information Density in R. P. Levy \& Jaeger (2007)
finds that the insertion of complementizers (e.g. ``that'') in relative
clauses in English corresponds to where neighboring words have high
information content. Similarly, A. F. Frank \& Jaeger (2008) argues that
contractions in English such as ``you're'' do not occur when neighboring
words are highly informative. Applications of Uniform Information
Density include determining whether linguistic alignment takes place
(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
2018) and the significance of ambiguity in language (Piantadosi, Tily,
\& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original work in R. P. Levy \& Jaeger
(2007), Zhan \& Levy (2018) focuses on information distribution at
particular points in sentences. Zhan \& Levy (2018) finds that more
information-rich classifiers in Mandarin Chinese are produced when
production of the neighboring noun is difficult, not when the
information content is high. Examining the information density at the
sentence level, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) study
word order across written sentences in Hindi, a freer word order
language than English, and find that information density has no
significant effect on determining a Hindi writer's word order.

@@

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found throughout English--in written language, and
in conversational speech between adults and between parents and their
children. Thus, this entropy curve is a robust feature of English
linguistic productions. We then examine entropy curves
cross-linguistically, and show that the characteristic entropy curves
vary across the world's languages. Finally, we show that this variation
is predictable in part from the structure of individual languages
(i.e.~word order). Taken together, our results suggest a refinement of
the Uniform Information Density hypothesis: speakers may structure their
utterances to optimize information density, but they must do so under
the predictable constraints of their language.

@@

With this measure, we can compute the unigram positional entropy at each
word position of sentences of each length within a corpus. The result of
this method across word positions can be plotted for each sentence
length as an \emph{entropy curve}, which can be visually compared across
sentence lengths to observe how the unigram entropy changes across
absolute positions in each of the sentences. Genzel \& Charniak (2002)
similarly examine a unigram entropy measure on sentences, and found that
entropy at the sentence level increases linearly with sentence index
within a corpus. Uniform Information Density applies this uniformity of
entropy rate in sentences to all levels of speech (R. P. Levy \& Jaeger,
2007), and so our method obtained from Yu et al. (2016), which examines
text at the word level, should find a monotonically increasing affine
function at the word level.

@@

We began by replicating Yu et al.'s (2016) analysis, computing entropy
curves on the British National Corpus--a collection of predominantly
written English (Clear, 1993). We then applied the same method to the
Switchboard corpus--a collection of spoken English (Godfrey, Holliman,
\& McDaniel, 1992). This allowed us to ask whether the characteristic
function identified by Yu et al. (2016) is a general feature of English,
or instead a function of their written language corpus.

\subsection{Data and Analysis}\label{data-and-analysis}

The British National Corpus consists of predominantly (90\%) written
documents collected in the early 1980s and 1990s across a variety of
genres (scientific articles, newspapers, fiction, etc). It also contains
a small collection of corpora from spoken language. All together, the
British National Corpus contains \(\sim\) 100 million tokens. The
Switchboard corpus is a collection of \(\sim\) 2,400 telephone
conversations between unacquainted adults prompted with subjects of
conversation. Switchboard is the corpus used in Levy and Jaeger's (2007)
original demonstration of the Uniform Information Density hypothesis.

@@

\subsection{Results}\label{results}

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lengths we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we found that the positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again for
the sentence-final word.

@@

\caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom). Points represent average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
\end{figure*}
\end{CodeChunk}

@@

Wikipedia has two primary advantages: (1) Individual language corpora
are generally large, and (2) hundreds of languages are represented. To
understand how variability in entropy curves is related to the structure
of language, we used features from the World Atlas of Language
Structures (WALS, Dryer \& Haspelmath, 2013). WALS is a curated
collection of typological features which help to classify the diversity
across the world's languages. However, as the collection includes
contributions from many authors studying a heterogeneous set of topics
and languages, most features are missing for most languages. For our
sample, we sought to balance diversity of languages with coverage in
WALS. We selected eight specific WALS features for which the same \(45\)
languages were coded (Figure \ref{fig:clust_tree}). Unfortunately, these
features all reflected the order of words in a sentence in each language
(e.g.~whether objects come before verbs, whether adjectives come before
nouns). A consistent set of results across these features would be
consistent with an effect of order on entropy curves, but unfortunately
no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In
order to understand how these curves varied across languages, we
developed a method for compressing the curves into a few representative
values. We noticed that curves tended to vary at predictable positions
within sentences: The majority were flat in the middle, but had either
rising or falling slopes at the beginnings and ends. For each language,
for each sentence length \(k\) we computed 5 slopes: The change between
the 1st and 2nd position, the change between the 2nd and 3rd position,
the slope between the 3rd position and the 3rd-to-last position
(\(k-1\)), the slope between the 3rd-to-last and the 2nd-to-last
position (\(k-1\)), and the slope between the 2nd-to-last and final
positions. We then averaged these slopes across sentence lengths within
a language to estimate a language-typical signature. We performed an
additional analysis dividing each sentence into fifths evenly and
computing the slopes between these fifths. Results from this second
analysis were qualitatively similar.

@@

We treated these five slopes as a vector, allowing us to compute the
pairwise similarity between languages with cosine--a standard measure of
vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
a smaller angle between them are more similar. Figure
\ref{fig:clust_tree} shows a hierarchical clustering of these 5
dimensional vectors, demonstrating that entropy curve similarity appears
to capture typological relatedly (e.g.~Slavic languages are similar to
each other).

We then asked whether these pairwise similarities were related to
overlap in linguistic features. For each of the eight WALS features we
measured, we fit a logistic regression predicting whether two languages
would have the same value for that feature from the pairwise similarity
of their entropy curves. Overlap on all eight features was reliably
predicted by entropy curve similarity (largest \(p = .001\)), with the
amount of variance in feature overlap predicted varying across
predictors \ref{tab:tab_features}. Order of object and verb (\(83A\))
was the strongest predictor, explaining 16.22\% of the variance in
pairwise entropy curve similarity.

@@

We expect the characteristic entropy curve for a language to have
important downstream effects for how people process that language. While
conversational turn-taking is universal, there exists cross-linguistic
variation in the average length of time between conversational partners
speaking (Stivers et al., 2009). We predict that one factor in this
variability relates to the shape of a language's entropy curve. A jump
at the end of a language's entropy curve, such as in English, indicates
that there is on average high levels of information at the end of
utterances in that language. Therefore we expect turn-taking times to be
longer in such languages than in languages where there is a drop in the
final segment of the entropy curve, such as Mandarin. We also expect
that readers of a language will take longer on average to read parts of
sentences where the entropy curve is relatively high, due to the
presence of higher information content. This may explain the so-called
wrap-up effect in sentence processing, in which the ends of sentences
are processed more slowly during reading (Kuperman, Dambacher, Nuthmann,
\& Kliegl, 2010; Stowe, Kaan, Sabourin, \& Taylor, 2018). The entropy
curve analysis can be carried out on any language across a wide variety
of text corpora, making it a powerful tool for analyzing the information
structure of both spoken and written utterances in a language.
